\section{Deductive Logic and Logical Connectives}
Deductive reasoning is the foundation on which proofs are based.
\ex{Deductive Reasoning}{
  \begin{enumerate}
    \item It will either rain or snow tomorrow.\\
          It's too warm to snow.\\
          Therefore, it will rain.
    \item If today is Sunday, then I don't have to work today.\\
          Today is Sunday.\\
          Therefore, I don't have to work today.
    \item I will work either tomorrow or today.\\
          I'm not working today.\\
          Therefore, I'm working tomorrow.
  \end{enumerate}
}
In each case we arrive at a conclusion given a set of premises which are assumed to be true.
The conclusion can only be false if at least one of the premises is false. 
If none of them are false, then the conclusion must be true.

\dfn{Valid Argument}{A valid argument is one in which the premises cannot all be true without the conclusion being true as well.}
\nt{An invalid argument is also known as a fallacy, in this case a false conclusion can be reached even if all the premises are true.}

\begin{table}[h]
  \centering
  \label{tab:logical_connectives}
  \begin{tabular}{|l|l|}
    \hline
    Natural Language & Symbol\\
    \hline
    And & \(\land\)\\
    Or & \(\lor\)\\
    Not & \(\neg\)\\
    If and only if & \(\iff\)\\
    Implies & \(\to\)\\
    Exclusive or & \(\oplus\)\\
    True & \(\top\)\\
    False & \(\bot\)\\
    For all & \(\forall\)\\
    There exists & \(\exists\)\\
    \hline
  \end{tabular}
  \caption{Logical Connectives}
\end{table}

\ex{Statements and their Logical Forms}{
  \begin{enumerate}
    \item Either John went to the store, or we're out of eggs.
    \item Joe is going to leave home and not come back.
    \item Either Bill is at work and Jane isn't, or Jane is at work and Bill isn't.
  \end{enumerate}
}

\newpage

\sol 
\begin{enumerate}
  \item $P \lor Q$ 
  \item $P \land Q$ 
  \item $(P \land \neg Q) \lor (\neg P \land Q)$ 
\end{enumerate}

\nt{
  $\neg$ always applies only to the statement immediately after it.
  $\neg P \land Q$ means  $(\neg P) \land Q$.
}

\input{chapters/chapter1/exercises-deductive.tex}
